---
title: "Lab 4c"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
librarian::shelf(
  digest, dplyr, DT, glue, purrr, readr, stringr, tidyr)

# path to folder containing species directories of images
dir_src  <- "/courses/EDS232/inaturalist-2021/train_mini"
dir_dest <- "~/inat"
dir.create(dir_dest, showWarnings = F)

# get list of directories, one per species (n = 10,000 species)
dirs_spp <- list.dirs(dir_src, recursive = F, full.names = T)
n_spp <- length(dirs_spp)

# set seed (for reproducible results) 
# just before sampling (otherwise get different results)
# based on your username (unique amongst class)
Sys.info()[["user"]] %>% 
  digest::digest2int() %>% 
  set.seed()
i10 <- sample(1:n_spp, 10)

# show the 10 indices sampled of the 10,000 possible 
i10
```

```{r}
# show the 10 species directory names
basename(dirs_spp)[i10]
```

```{r}
# show the first 2 species directory names
i2 <- i10[1:2]
basename(dirs_spp)[i2]
```

```{r}
# setup data frame with source (src) and destination (dest) paths to images
d <- tibble(
  set     = c(rep("spp2", 2), rep("spp10", 10)),
  dir_sp  = c(dirs_spp[i2], dirs_spp[i10]),
  tbl_img = map(dir_sp, function(dir_sp){
    tibble(
      src_img = list.files(dir_sp, full.names = T),
      subset  = c(rep("train", 30), rep("validation", 10), rep("test", 10))) })) %>% 
  unnest(tbl_img) %>% 
  mutate(
    sp       = basename(dir_sp),
    img      = basename(src_img),
    dest_img = glue("{dir_dest}/{set}/{subset}/{sp}/{img}"))

# show source and destination for first 10 rows of tibble
d %>% 
  select(src_img, dest_img)
```

```{r}
# iterate over rows, creating directory if needed and copying files 
d %>% 
  pwalk(function(src_img, dest_img, ...){
    dir.create(dirname(dest_img), recursive = T, showWarnings = F)
    file.copy(src_img, dest_img) })

# uncomment to show the entire tree of your destination directory
# system(glue("tree {dir_dest}"))
```

### Task 1: 2 Species (Binary Classification) - nerual net

```{r}
library(keras)
```

Create directories to access images

```{r}
spp2_train_dir <- paste0(dir_dest,"/spp2/train")
spp2_test_dir <- paste0(dir_dest,"/spp2/test")
spp2_validation_dir <- paste0(dir_dest, "/spp2/validation")
```

```{r}
# Re-scale all images by 1/255
train_datagen <- image_data_generator(rescale = 1/255)
validation_datagen <- image_data_generator(rescale = 1/255)
test_datagen <- image_data_generator(rescale = 1/255)
```

```{r}
spp2_train_generator <- flow_images_from_directory(
  spp2_train_dir,
  train_datagen,
  target_size = c(150, 150),
  batch_size = 15,
  class_mode = "binary"
)

spp2_validation_generator <- flow_images_from_directory(
  spp2_validation_dir,
  validation_datagen,
  target_size = c(150, 150),
  batch_size = 15,
  class_mode = "binary"
)

spp2_test_generator <- flow_images_from_directory(
  spp2_test_dir,
  test_datagen,
  target_size = c(150, 150),
  batch_size = 15,
  class_mode = "binary"
)
```

```{r}
spp2_model <- keras_model_sequential() %>% 
  layer_dense(units = 16, activation = "relu", input_shape = c(150, 150, 3)) %>% 
  layer_dense(units = 16, activation = "relu") %>% 
  layer_flatten() %>% 
  layer_dense(units =  1, activation = "sigmoid")
```

```{r}
spp2_model %>% compile(
  optimizer = optimizer_rmsprop(learning_rate = 0.001),
  loss = "binary_crossentropy",
  metrics = c("accuracy")
)
```

```{r}
summary(spp2_model)
```

```{r}
spp2_history <- spp2_model %>% fit(
    spp2_train_generator,
    steps_per_epoch = 1,
    epochs = 20,
    validation_data = spp2_validation_generator,
    validation_steps = 1
  )
```

```{r}
plot(spp2_history)
```

```{r}
spp2_model %>% predict(spp2_test_generator)

spp2_model %>% evaluate(spp2_test_generator)
```

### Task 2: 2 Species (binary classification) - convolutional neural net

```{r}
convnet_model <- keras_model_sequential() %>% 
  layer_conv_2d(filters = 32, kernel_size = c(3, 3), activation = "relu",
                input_shape = c(150, 150, 3)) %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_flatten() %>% 
  layer_dense(units = 512, activation = "relu") %>% 
  layer_dense(units = 1, activation = "sigmoid")
```

```{r}
summary(convnet_model)
```

```{r}
convnet_model %>% compile(
  loss = "binary_crossentropy",
  optimizer = optimizer_rmsprop(learning_rate = 0.0001),
  metrics = c("acc"))
```

```{r}
convnet_model_history <- convnet_model %>% fit(
    spp2_train_generator,
    steps_per_epoch = 1,
    epochs = 30,
    validation_data = spp2_validation_generator,
    validation_steps = 1
)

str(convnet_model_history)
```

```{r}
plot(convnet_model_history)

convnet_model %>% predict(spp2_test_generator)

convnet_model %>% evaluate(spp2_test_generator)
```

#### Comparing the two species binary models:

**For the 2 species binary models, the convnets model seems to give a higher accuracy with a smaller amount of loss than the regular binary classification model, so I would go with that model in this situation.**

### Task 3: 10 Species (multi-class classification) - neural net

Create new directories for 10 species

```{r}
spp10_train_dir <- paste0(dir_dest,"/spp10/train")
spp10_test_dir <- paste0(dir_dest,"/spp10/test")
spp10_validation_dir <- paste0(dir_dest, "/spp10/validation")
```

```{r}
spp10_train_generator <- flow_images_from_directory(
  spp10_train_dir,
  train_datagen,
  target_size = c(150, 150),
  batch_size = 10,
  class_mode = "categorical"
)

spp10_validation_generator <- flow_images_from_directory(
  spp10_validation_dir,
  validation_datagen,
  target_size = c(150, 150),
  batch_size = 10,
  class_mode = "categorical"
)

spp10_test_generator <- flow_images_from_directory(
  spp10_test_dir,
  test_datagen,
  target_size = c(150, 150),
  batch_size = 10,
  class_mode = "categorical"
)
```

```{r}
multi_class_model <- keras_model_sequential() %>% 
  layer_dense(units = 64, activation = "relu", input_shape = c(150, 150, 3)) %>% 
  layer_dense(units = 64, activation = "relu") %>% 
  layer_flatten() %>% 
  layer_dense(units = 10, activation = "softmax")
```

```{r}
multi_class_model %>% compile(
  optimizer = "rmsprop",
  loss = "categorical_crossentropy",
  metrics = c("acc")
)
```

```{r}
spp10_history <- multi_class_model %>% fit(
  spp10_train_generator,
  steps_per_epoch = 5,
  epochs = 20,
  validation_data = spp10_validation_generator,
  validation_steps = 10
  )
```

```{r}
plot(spp10_history)
```

```{r}
multi_class_model %>% predict(spp10_test_generator)

multi_class_model %>% evaluate(spp10_test_generator)
```

### Task 4: 10 Species (multi-class classification) - convolutional neural net

```{r}
multi_class_convnet_model <- keras_model_sequential() %>% 
  layer_conv_2d(
    filters = 32, kernel_size = c(3, 3), activation = "relu",
    input_shape = c(150, 150, 3)) %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 64, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_conv_2d(filters = 128, kernel_size = c(3, 3), activation = "relu") %>% 
  layer_max_pooling_2d(pool_size = c(2, 2)) %>% 
  layer_flatten() %>% 
  layer_dropout(rate = 0.5) %>% 
  layer_dense(units = 512, activation = "relu") %>% 
  layer_dense(units = 10, activation = "sigmoid") 
```

```{r}
multi_class_convnet_model %>% compile(
  loss = "categorical_crossentropy",
  optimizer = optimizer_rmsprop(learning_rate = 1e-4),
  metrics = c("acc"))
```

```{r}
spp10_history_convnet <- multi_class_convnet_model %>% fit(
    spp10_train_generator,
    steps_per_epoch = 5,
    epochs = 30,
    validation_data = spp10_validation_generator,
    validation_steps = 10
)
```

```{r}
plot(spp10_history_convnet)
```

```{r}
multi_class_convnet_model %>% predict(spp10_test_generator)

multi_class_convnet_model %>% evaluate(spp10_test_generator)
```

#### Comparing the ten species multi-class models:

**For the 10 species multi-class models, the regular multi-class model seems to give a higher accuracy with a smaller amount of loss than the convnet multi-class classification model, so I would go with that model in this situation. However, here both models seem to be pretty low accuracy which leads me to believe that you would either want to build a model with more data, or do a better job of tuning the parameters than I did.**


